---
title: "Homework 2"
author: "Keyanna Davis"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


#Problem 1

```{r Problem 1}
mr_trash_wheel = readxl::read_excel("./data/trash_wheel.xlsx", 
                                sheet=1)%>%
  janitor::clean_names() %>% 
  select(-x15,-x16,-x17) %>% 
  drop_na() %>% 
mutate(
  sports_balls=round(sports_balls),
  sports_balls=as.integer(sports_balls)
)
    

 
pdata_17 = readxl::read_excel("./data/trash_wheel.xlsx",
                                    sheet=6, skip = 1) %>% 
janitor::clean_names() %>% 
drop_na() %>% 
  mutate(.data=., year="2017"
) 


pdata_18 = readxl::read_excel("./data/trash_wheel.xlsx",
                                    sheet=5, skip=1) %>% 
janitor::clean_names() %>% 
drop_na() %>% 
  mutate(.data=., year="2018"
) 


pdata = full_join(pdata_17, pdata_18, by= c("month"="month", "year"="year",
                                            "total"="total")) %>% 
  mutate(month = month.name[month] )
    
```

###Summary of Problem 1

The number of observations in Mr. Trash Wheel dataset is `r nrow(mr_trash_wheel)` and `r ncol(mr_trash_wheel)` variables, 2017 Precipation dataset is `r nrow(pdata_17)` observations and `r ncol(pdata_17)` variables, 2018 Precipation dataset is `r nrow(pdata_18)` observations and `r ncol(pdata_18)` variables. The 2017 and 2018 Precipation dataset combined has `r nrow(pdata)` observations and `r ncol(pdata)` variables.



#Problem 2

```{r}

pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
month = as.numeric(month),
month = month.abb[month] ) %>% 
  mutate( 
  president = recode(prez_gop,
          '0' = "republican" ,
          '1' = "democratic")
  ) %>% 

  select(-prez_dem, -prez_gop,-day)



snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>% 
  mutate(
month = as.numeric(month),
month = month.abb[month] )%>% 
  select(year, month, everything())


unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
    pivot_longer(
      jan:dec,
      names_to="month",
      values_to = "day"
    ) %>% 
  mutate(
     day = round(day),
     day = as.integer(day)
  )

pols_snp = full_join(pols_month, snp, by = c("year"="year", "month"="month")) %>% 
  mutate(
    year=as.numeric(year) 
    )%>% 
    select(-day)
    
  

pols_snp_unemploy = full_join(pols_snp, unemployment, by=c( "month"="month", "year"="year")) %>% 
  select(-day) %>% 
  drop_na()

```



###Summary of Problem 2




#Problem 3

```{r problem 3}
baby_names = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  
  mutate(
   ethnicity = recode(ethnicity, 
           "ASIAN AND PACIFIC ISLANDER"="ASIAN AND PACI",
           "BLACK NON HISPANIC" = "BLACK NON HISP",
           "WHITE NON HISPANIC" = "WHITE NON HISP")
  ) %>% 
  mutate(
    childs_first_name = str_to_upper(childs_first_name)
  ) %>% 
distinct()


 new_df = filter(baby_names, gender=="FEMALE", childs_first_name=="OLIVIA") %>% 
   select(-count)

 popular_girl =
 new_df %>% 
   select(rank,ethnicity, year_of_birth) %>% 
  pivot_wider(
   names_from = year_of_birth,
   values_from = rank
 )
  popular_girl
  
  
  
  male = filter(baby_names, gender=="MALE", count == max(count)) %>% 
  select(gender,count,childs_first_name, rank)
  
   new_df2 = filter(baby_names, gender=="MALE", childs_first_name=="JAYDEN") %>% 
   select(-count)
  
    popular_guy =
 new_df2 %>% 
   select(rank,ethnicity, year_of_birth) %>% 
  pivot_wider(
   names_from = year_of_birth,
   values_from = rank
 )
popular_guy



new_df3 = filter(baby_names, gender=="MALE", ethnicity=="WHITE NON HISP", year_of_birth=="2016") 


scatter_plot = ggplot(new_df3, aes(y=count, x=rank))+
  geom_point()
scatter_plot
```




