---
title: "Homework 2"
author: "Keyanna Davis"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```


#Problem 1

```{r Problem 1}
mr_trash_wheel = readxl::read_excel("./data/trash_wheel.xlsx", 
                                sheet=1)%>%
  janitor::clean_names() %>% 
  select(-x15,-x16,-x17) %>% 
  drop_na() %>% 
mutate(
  sports_balls=round(sports_balls),
  sports_balls=as.integer(sports_balls)
)
    

 
pdata_17 = readxl::read_excel("./data/trash_wheel.xlsx",
                                    sheet=6, skip = 1) %>% 
janitor::clean_names() %>% 
drop_na() %>% 
  mutate(.data=., year="2017"
) 


pdata_18 = readxl::read_excel("./data/trash_wheel.xlsx",
                                    sheet=5, skip=1) %>% 
janitor::clean_names() %>% 
drop_na() %>% 
  mutate(.data=., year="2018"
) 


pdata = full_join(pdata_17, pdata_18, by= c("month"="month", "year"="year",
                                            "total"="total")) %>% 
  mutate(month = month.name[month] )
   



sports_balls_2017 = filter(mr_trash_wheel, year == "2017", sports_balls)
median(pull(sports_balls_2017, sports_balls))
```

####Summary of Problem 1

The number of observations in`Mr. Trash Wheel` dataset has `r nrow(mr_trash_wheel)` and `r ncol(mr_trash_wheel)` variables. The average of plastic bottles in `Mr. Trash Wheel` is `r mean(pull(mr_trash_wheel, plastic_bottles))`, and the average of cigarette butts is `r mean(pull(mr_trash_wheel, cigarette_butts))` which is fairly a lot of cigarette butts. Also in `Mr. Trash Wheel` dataset the median of sports ball of year 2017 is `r median(pull(sports_balls_2017, sports_balls)) `.


In `2017 Precipation` dataset has `r nrow(pdata_17)` observations and `r ncol(pdata_17)` variables. The `total precipation of year 2017` is `r sum(pull(pdata_17, total))`. In `2018 Precipation` dataset has `r nrow(pdata_18)` observations and `r ncol(pdata_18)` variables. The `total precipation of year 2018` is `r sum(pull(pdata_18, total))`. 


The `2017 and 2018 Precipation` dataset combined has `r nrow(pdata)` observations and `r ncol(pdata)` variables. The `total precipation of year 2017 and 2018` combined is `r sum(pull(pdata, total))`. 





#Problem 2

```{r}

pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(
month = as.numeric(month),
month = month.abb[month] ) %>% 
  mutate( 
  president = recode(prez_gop,
          '0' = "republican" ,
          '1' = "democratic")
  ) %>% 

  select(-prez_dem, -prez_gop,-day)



snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>%
     mutate(month = as.numeric(month)) %>% 
  arrange(year,month) %>% 
  mutate(
   month = month.abb[month] 
  ) %>% 
  
  
  select(year, month, everything())


unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
    pivot_longer(
      jan:dec,
      names_to="month",
      values_to = "day"
    ) %>% 
  mutate(
     day = round(day),
     day = as.integer(day)
  )

pols_snp = left_join(pols_month, snp, by = c("year"="year", "month"="month")) %>% 
  mutate(
    year=as.numeric(year) 
    )%>% 
    select(-day)
    
  

pols_snp_unemploy = left_join(pols_snp, unemployment, by=c( "month"="month", "year"="year")) %>% 
  select(-day) %>% 
  drop_na()

```



####Summary of Problem 2
The above datasets are ` pols_month, snp, unemployment` and `pols_snp_unemply`, which contains the three datasets combined. The number of observations in `pols_month` dataset has `r nrow(pols_month)` and `r ncol(pols_month)` variables. The years in `pols_month` ranges from `r range(pull(pols_month,year))`. Some key variables in the dataset are `month` and `year`. 


In the `snp` dataset their are `r nrow(snp)` observations and `r ncol(snp)` variables. The years in `pols_month` ranges from `r range(pull(snp,year))`. There are some key variables in `snp` dataset which include `month` and `year`.


In the `unemployment` dataset their are `r nrow(unemployment)` observations and `r ncol(unemployment)` variables. The years in `unemployment` ranges from `r range(pull(unemployment,year))`. Some key variables in the dataset are `month` and `year`. 


In the combined dataset `pols_snp_unemploy` their are `r nrow(pols_snp_unemploy)` observations and `r ncol(pols_snp_unemploy)` variables. This dataset `pols_snp_unemploy` years ranges from `r range(pull(pols_snp_unemploy, year))`. The key variables here are also `month` and `year`.




#Problem 3

```{r problem 3}
baby_names = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  
  mutate(
   ethnicity = recode(ethnicity, 
           "ASIAN AND PACIFIC ISLANDER"="ASIAN AND PACI",
           "BLACK NON HISPANIC" = "BLACK NON HISP",
           "WHITE NON HISPANIC" = "WHITE NON HISP")
  ) %>% 
  mutate(
    childs_first_name = str_to_upper(childs_first_name)
  ) %>% 
distinct()


 new_df = filter(baby_names, gender=="FEMALE", childs_first_name=="OLIVIA") %>% 
   select(-count)

 popular_girl =
 new_df %>% 
   select(rank,ethnicity, year_of_birth) %>% 
  pivot_wider(
   names_from = year_of_birth,
   values_from = rank
 )
 
popular_girl %>% 
   knitr::kable()
  
  
  
  male = filter(baby_names, gender=="MALE", rank == min(rank)) %>% 
  select(gender,count,childs_first_name, rank)
  
   new_df2 = filter(baby_names, gender=="MALE", childs_first_name=="ETHAN") %>% 
   select(-count)
  
    popular_guy =
 new_df2 %>% 
   select(rank,ethnicity, year_of_birth) %>% 
  pivot_wider(
   names_from = year_of_birth,
   values_from = rank
 ) 
 popular_guy %>%   
knitr::kable()


new_df3 = filter(baby_names, gender=="MALE", ethnicity=="WHITE NON HISP", year_of_birth=="2016") 


scatter_plot = ggplot(new_df3, aes(y=count, x=rank ))+
  geom_point()

scatter_plot
```




